---
title: "Case Study - Abdominal aortic aneurysm"
author: "Xavier G.L.V. Pouwels"
format: pdf
editor: visual
bibliography: references.bib
---

# Introduction

This vignette showcases the functionalities of `pacheck` on the [AAA_DES_model](https://github.com/mikesweeting/AAA_DES_model), which is described in @sweeting2018 . The probabilistic set of inputs and outputs has been obtained by running the `Main file_men 30years_FullModel.R` file under the `models` folder of the R project. To limit computational time, we only simulated 1,000 iterations containing 1,000 individuals (which may not be enough to obtain stable results).

# Loading packages and data

```{r}
library(pacheck)
library(purrr)
data("l_psa_aaa")
set.seed(1234)
```

# Preparing dataset for use in `pacheck`

The following code chunk prepares the inputs and outputs list of the AAA DES model to use the `pacheck` tests. The examples included in this appendix are performed on a subset of all model inputs and outputs to limit the size of the outputs provided by the `pacheck` tests. The functionalities of `pacheck` still work on a larger number of HE model inputs since the tests included in `pacheck` are performed on the column-level of the dataframe of inputs and outputs. The number of columns/ parameters included in the check therefore does not affect the working of the tests implemented in `pacheck`.

```{r}
# Inputs
## Input parameters are stored in a list and 
##have attributes that are used here to retrieve 
##the different types of parameters
##using the `map_depth` function from the `purrr` package
v_attr_type <- unlist(lapply(l_psa_aaa$v2values[[1]], function(l_obj)
  attr(l_obj, which = "type")), use.names = F)
v_probnames <- names(l_psa_aaa$v2value[[1]])[which(v_attr_type == "probability")]
v_ratenames <- names(l_psa_aaa$v2value[[1]])[grep("rate", v_attr_type)]
v_qalynames <- names(l_psa_aaa$v2value[[1]])[grep("qaly", v_attr_type)]
v_costnames <- names(l_psa_aaa$v2value[[1]])[grep("cost", v_attr_type)]
v_parmodnames <- names(l_psa_aaa$v2value[[1]])[grep("par", v_attr_type)]
v_fixed <- names(l_psa_aaa$v2value[[1]])[grep("fixed", v_attr_type)]
m_aaa_inputs <- sapply(c(v_probnames, 
                         v_ratenames, 
                         v_fixed,
                         # v_qalynames, # omitted because fixed
                         # v_costnames, # omitted because fixed
                         v_parmodnames), function(name)
                           unlist(purrr::map_depth(l_psa_aaa$v2values, 1, name),
                                  use.names = F))
# Outputs
## Outputs are stored in the 3D array
m_aaa_res <- l_psa_aaa$psaQuantities
dim(m_aaa_res) <- c(unname(
  dim(l_psa_aaa$psaQuantities)[[1]] * dim(l_psa_aaa$psaQuantities)[[2]]
), unname(dim(l_psa_aaa$psaQuantities)[[3]]))
m_aaa_res <- t(m_aaa_res)
colnames(m_aaa_res) <- paste0(unlist(dimnames(l_psa_aaa$psaQuantities)[1]), 
                              "_", 
                              rep(unlist(dimnames(l_psa_aaa$psaQuantities)[2]),
                                  each = length(unlist(
                                    dimnames(l_psa_aaa$psaQuantities)[1]))))
df_aaa <- data.frame(cbind(m_aaa_inputs, m_aaa_res))
```

# Testing plausibility of model inputs and outputs

In this section, we test the plausibility of multiple groups of parameters and outcomes using the `do_quick_check` function. We also check whether discounted results are lower than corresponding non-discounted results using the `do_discount_check` function.

```{r}
# Quick checks of groups of input parameter values
do_quick_check(df = df_aaa,
               v_probs = v_probnames,
               v_r = v_ratenames,
               v_costs = c("noScreening_cost",
                           "screening_cost",
                           "noScreening_discountedCost",
                           "screening_discountedCost"))

# Discounted results lower than undiscounted results
do_discount_check(
  df = df_aaa,
  v_outcomes = c(
    "noScreening_lifeYears",
    "screening_lifeYears",
    "noScreening_qalys",
    "screening_qalys",
    "noScreening_cost",
    "screening_cost"
  ),
  v_outcomes_d = c(
    "noScreening_discountedLifeYears",
    "screening_discountedLifeYears",
    "noScreening_discountedQalys",
    "screening_discountedQalys",
    "noScreening_discountedCost",
    "screening_discountedCost"
  )
)
```

# Drawing health economic figures

`pacheck` also contains figures to draw standard figures for health economic evaluations, such as the cost-effectiveness plane (`plot_ce` function), the incremental cost-effectiveness plane (`plot_ice` function), the cost-effectiveness acceptability curves (`plot_ceac` function), and net benefit curves (`plot_nb` function). Both the `plot_ceac` and `plot_nb` functions require data frames containing specific variable names, which can respectively be obtained with the accompanying functions `calculate_ceac` and `calculate_nb`. The clouds of incremental costs and effects plotted incremental cost-effectiveness plane below does not show the typical ellipsoidal shape that is observed in most health economic analysis. Hence, this observation would require a deeper investigation of the health economic model's inputs and outputs. In this case, the shape may be cause by a model input parameter taking (almost) discrete values.

```{r}
plot_ce(
  df = df_aaa,
  e_int = "screening_discountedQalys",
  e_comp = "noScreening_discountedQalys",
  c_int = "screening_discountedCost",
  c_comp = "noScreening_discountedCost",
  axes = F
)
plot_ice(
  df = df_aaa,
  e_int = "screening_discountedQalys",
  e_comp = "noScreening_discountedQalys",
  c_int = "screening_discountedCost",
  c_comp = "noScreening_discountedCost"
)
df_ceac <- calculate_ceac(
  df = df_aaa,
  e_int = "screening_discountedQalys",
  e_comp = "noScreening_discountedQalys",
  c_int = "screening_discountedCost",
  c_comp = "noScreening_discountedCost"
)  
plot_ceac(df = df_ceac, name_wtp = "WTP_threshold")
df_nb <- calculate_nb(
  df = df_aaa,
  e_int = "screening_discountedQalys",
  e_comp = "noScreening_discountedQalys",
  c_int = "screening_discountedCost",
  c_comp = "noScreening_discountedCost",
  wtp = 80000
)
plot_nb(df = df_nb)
```

# Reference
