---
title: "Case study - iviRA model"
author: "Xavier G.L.V. Pouwels"
format: pdf
editor: visual
bibliography: references.bib
---

# Introduction

This vignette shows how to use the functionalities of `pacheck` on the `iviRA` health economic model [@incerti2019].

# Loading packages and data

```{r}
library(pacheck)
set.seed(1234)
data("df_iviRA_pa")
```

# Generating the probabilistic inputs and outputs

The source code to generate the model inputs and outputs of the `iviRA` cost-effectiveness model is available in the source code of the `pacheck` package on github (<https://github.com/Xa4P/pacheck/blob/master/analysis/03-iviRA_data_generation.R>). This script converts the probabilistic inputs and outputs list of a health economic analysis performed using the `iviRA` package to a dataframe of probabilistic inputs and outputs on which the `pacheck` tests can be performed.

The examples included in this appendix are performed on a subset of all model inputs and outputs to limit the size of the outputs provided by the `pacheck` tests. The functionalities of `pacheck` still work on a larger number of health economic (HE) model inputs since the tests included in `pacheck` are performed on the column-level of the dataframe of inputs and outputs. The number of columns/ parameters included in the check therefore does not affect the working of the tests implemented in `pacheck`.

# Testing plausibility of model inputs and outputs

One way to test the plausibility of model inputs and outputs is simply to inspect the summary statistics of each variable. This can be done with the `pacheck` package using the `generate_sum_stats` function.

```{r}
# Summary stats (10 first model inputs for long-term HAQ progression)
generate_sum_stats(df = df_iviRA_pa[, grep("r_haq_prog", 
                                           names(df_iviRA_pa))])[1:10,] 

# Specific checks
## Proportions & probabilities between 0 and 1
### p_discount parameters are discounts (rebates) on the prices of the medicines
v_vars_discount <- grep("p_discount", names(df_iviRA_pa), value = TRUE) # returns
# variable names containing p_discount
check_binary(c(v_vars_discount, "p_rebound"), df = df_iviRA_pa)

## Costs, counts, OR positive
v_vars_n <- grep("n_", names(df_iviRA_pa), value = TRUE)
v_vars_c <- grep("c_", names(df_iviRA_pa), value = TRUE)
v_vars_c <- v_vars_c[-grep("inc_", v_vars_c)]
check_positive(c(v_vars_n, v_vars_c, "or_mort"), df = df_iviRA_pa)
do_check(df = df_iviRA_pa,
         v_vars = c(v_vars_n, v_vars_c, "or_mort"),
         check = ~ .x >= 0,
         label_check = "positive") # alternative
```

# Metamodelling

In this section, we will perform metamodelling using two techniques: linear regression modelling (lm) and random forest (rf) using all input parameters of the probabilistic analysis data frame. Before performing metamodelling, we will standardise the inputs using their respective mean and standard deviation.

## Fit

The metamodels were first developed on the full data sets, using all input parameters. The dependent variable of these metamodels is the "incremental QALY" between the two strategies we evaluated using the health economic model implemented in `iviRA`. Note that wew did not perform variable selection, which could be performed to improve the metamodels' fit. For the sake of simplicity, we will use the same input variables for the linear regression metamodel and the random forest metamodel.

In the `fit_rf_metamodel`, we set the `var_importance` argument to `TRUE` to be able to extract variable importance plots and for the random forest metamodel.

```{r}
# Select parameters for metamodelling
v_x_vars <- names(df_iviRA_pa)
v_x_vars <- v_x_vars[-grep("t_qaly", v_x_vars)]
v_x_vars <- v_x_vars[-grep("t_cost", v_x_vars)]
v_x_vars <- v_x_vars[-grep("inc_cost", v_x_vars)]
v_x_vars <- v_x_vars[-grep("inc_qaly", v_x_vars)]
v_x_vars <- v_x_vars[-grep("p_discount", v_x_vars)]
v_x_vars <- v_x_vars[-grep("^c_", v_x_vars)]
#remove cost-related parameters

## standardisation
df_iviRA_mm <- df_iviRA_pa
df_iviRA_mm[, v_x_vars] <- as.data.frame(apply(df_iviRA_mm[, v_x_vars], 
                                               2, 
                                               function(col) {
                                                 (col - mean(col)) / sd(col)}
                                               )
                                         )

# Fit linear regression metamodel using all inputs contained in the dataframe
metamodel_lm <- fit_lm_metamodel(df = df_iviRA_mm,
                                 y_var = "inc_qaly",
                                 x_vars = v_x_vars)
summary(metamodel_lm$fit)
y <- df_iviRA_mm[, metamodel_lm$model_info$y_var]
predicted <- metamodel_lm$fit$fitted.values
r_squared_lm <- 1 - sum((y-predicted)^2)/sum((y-mean(y))^2)

metamodel_rf <- fit_rf_metamodel(df = df_iviRA_mm,
                                 y_var = "inc_qaly",
                                 x_vars = v_x_vars,
                                 ntree = 1000,
                                 var_importance = TRUE # plots variable importance
                                 ) 
# R-squared for RF model because not exported by model summary
y <- metamodel_rf$fit$yvar
predicted <- metamodel_rf$fit$predicted
r_squared_rf <- 1 - sum((y-predicted)^2)/sum((y-mean(y))^2)
```

### Interpretation

The R^2^ of the linear regression metamodel, `r round(r_squared_lm, 3)` is lower for this HE model than for the HE model implemented in the `cdx2cea` package. This can partly be explained by the patient-level nature of the HE model implemented in `iviRA` which may result in a less linear relationship between HE model inputs and outputs. In contrast, the R^2^ of the random forest metamodel is `r round(r_squared_rf, 3)`, which is substantially higher than the R^2^ of the linear regression metamodel. These two R^2^ estimates indicate how well the both metamodel fit the training data, they however do not indicate how accurate the models are at predicting outcomes in data that is not used to train the metamodels. This is something we will investigate under the validation subheading.

Concerning the random forest metamodel, one can see that the error rate is not decreasing anymore once 1,000 are used in the random forest algorithm. This suggests that investigating the use of more trees within the algorithm may not lead to a further reduction of the error rate. Based on the values of the coefficients of the linear regression metamodels and the variable importance plots of the random forest metamodel (second plot), one can conclude that the model parameters relating to the progression of the HAQ score have the highest impact on the incremental QALYs.

\newpage

## Validation

Validation of the metamodel will be done using cross-validation with ten folds.

```{r}
# Metamodels validation
validation_mm_lm <- validate_metamodel(
  model = metamodel_lm,
  method = 'cross_validation',
  folds = 10
)
validation_mm_rf <- validate_metamodel(
  model = metamodel_rf,
  method = 'cross_validation',
  folds = 10
)
knitr::kable(validation_mm_lm$stats_validation, 
             caption = "Validation metrics of the linear regression metamodel.")
knitr::kable(validation_mm_rf$stats_validation, 
             caption = "Validation metrics of the random forest metamodel with 1,000 trees.")
```

### Interpretation

Based on this validation step, both metamodels result in low R^2^ estimates and high mean absolute and relative errors in the training sets. These high R^2^ in the training sets but low R^2^ in the validation sets may indicate overfitting of the metamodels in the training set and a low generalisability of the metamodels. Hence, both models should actually not be used for making predictions. We will however perform some predictions using these metamodels only for illustration purposes.

## Prediction

The following code snippet shows how to make prediction using the fitted metamodels.

```{r}
# Create dataframe containing the 10th and 90th percentile of the probabilistic values of the seleted variables in the metamodels
df_inputs_pred <- data.frame(matrix(ncol = length(v_x_vars),
                             nrow = 2,
                             dimnames = list(c(),
                                             v_x_vars))
                      )
df_inputs_pred[1, ] <- unname(apply(df_iviRA_mm[, v_x_vars], 
                             2, 
                             function(rows) {quantile(rows, probs = 0.1)}))
df_inputs_pred[2, ] <- unname(apply(df_iviRA_mm[, v_x_vars], 
                             2, 
                             function(rows) {quantile(rows, probs = 0.9)}))

# Predict
df_lm_preds <- predict_metamodel(model = metamodel_lm,
                                 inputs = df_inputs_pred,
                                 output_type = "long_df")
df_rf_preds <- predict_metamodel(model = metamodel_rf,
                                 inputs = df_inputs_pred,
                                 output_type = "long_df")

# Display
knitr::kable(df_lm_preds, 
             caption = "Input parameter values and predictions using the linear regression metamodel.")
knitr::kable(df_rf_preds, 
             caption = "Input parameter values and predictions using the random forest metamodel.")
```

# References
