---
title: "Case study - Screening CKD CEA"
author: "Xavier G.L.V. Pouwels"
format: pdf
editor: visual
bibliography: references.bib
---

# Introduction

This vignette showcases how to use the functionalities of `pacheck` on a health economic model developed in Python focusing on the health economic impact on screening for Chronic Kidney Disease [@cusick2023].

# Loading packages and data

```{r}
library(pacheck)
set.seed(1234)
data("df_ckd_inputs")
data("df_ckd_results")
```

# Preparing dataset for use in `pacheck`

The following code chunk prepares the inputs and outputs list of the CKD Screening CEA to use the `pacheck` tests. In this example, only a subset of all inputs is used to limit the size of the output provided by `pacheck`.

```{r, warning = FALSE}
df_ckd_results_wide <- df_ckd_results |>
  mutate(intervention = ifelse(intervention == "[0, 10, 20, 30, 40]", 
                               "Screening", 
                               intervention)) |>
  select(!c(full_index, Starting.age, dapa.costs, avg.on.dapa)) |>
  pivot_wider(names_from  = intervention,
              names_sep = "_",
              values_from = !intervention
              )
df_ckd_results_wide <- data.frame(apply(df_ckd_results_wide, 2, unlist))
df_ckd <- cbind(df_ckd_inputs,
                df_ckd_results_wide)
df_ckd$Unnamed..0 <- df_ckd$iteration <- NULL

# Add utility values defined in model script
## General population utility values
df_ckd$male_35_QALY <- 0.925
df_ckd$male_45_QALY <- 0.894 
df_ckd$male_55_QALY <- 0.870 
df_ckd$male_65_QALY <- 0.852
df_ckd$male_75_QALY <- 0.816
df_ckd$male_85_QALY <- 0.807

df_ckd$female_35_QALY <- 0.900
df_ckd$female_45_QALY <- 0.871 
df_ckd$female_55_QALY <- 0.846 
df_ckd$female_65_QALY <- 0.822
df_ckd$female_75_QALY <- 0.784
df_ckd$female_85_QALY <- 0.747
```

# Testing plausibility of model inputs and outputs

One way to test the plausibility of model inputs and outputs is simply to inspect the summary statistics of each variable. This can be done with the `pacheck` package using the `generate_sum_stats` function. The correlation between (specific) inputs and outputs of the health economic model can be generated using the `generate_cor` function. This allows to investigate the direction and strengths of the relationships between HE model inputs and outputs.

In addition, one can check whether the mean quality of life of individuals within the model (total undiscounted quality adjusted life years divided by total undiscounted life years) lies within the minimal and maximal utility value in each iteration (using the `check_mean_qol` function).

```{r}
# Summary stats
generate_sum_stats(df = df_ckd[, c(grep("Screening", names(df_ckd)),
                                   grep("Placebo", names(df_ckd)))]) 
## only for the outcomes for the sake of brevety

# Check correlations between inputs and outputs
##e.g. screening costs and higher ACE.costs should lead to higher screening costs and total costs
generate_cor(df = df_ckd,
             vars = c(
               "screening_cost",
               "ACE.cost",
               "ace.costs_Screening",
               "ace.costs_Placebo",
               "costs_Screening",
               "costs_Placebo"
             ),
             digits = 3)
## weak negative correlation between screening_cost and costs_Screening, 
## weak positive correlation between ACE.cost and total ACE costs 
## weak negative correlation with total costs of strategies 

# Specific checks
## mean quality of life between minimal and maximal utility values
v_vars_u <- grep("_QALY", names(df_ckd), value = TRUE) 
v_vars_u <- v_vars_u[-grep("dis", v_vars_u)]

check_mean_qol(df = df_ckd,
               t_qaly = "QALY_Screening",
               t_ly = "LY_Screening",
               u_values = v_vars_u)

check_mean_qol(df = df_ckd,
               t_qaly = "QALY_Placebo",
               t_ly = "LY_Placebo",
               u_values = v_vars_u)

## Costs, prevalence, HR are positive
v_vars_prev <- grep("prev", names(df_ckd), value = TRUE)
v_vars_c <- grep("_cost", names(df_ckd), value = TRUE)
v_vars_hr <- grep(".HR", names(df_ckd), value = TRUE)
# v_vars_c <- v_vars_c[-grep("inc_", v_vars_c)]
check_positive(c(v_vars_prev, v_vars_c, v_vars_hr), df = df_ckd)
```

# Reference
