---
title: "Do-It-Yourself model validation: a practical illustration of the Probabilistic Analysis Check R Package in four case studies"
author: "X.G.L.V. Pouwels & H. Koffijberg"
date: "`r Sys.Date()`"
output: word_document
bibliography: references.bib
csl: pharmacoeconomics.csl
---

```{css, echo=FALSE}
.box-style {
  background-color: white;
  border: 3px solid black;
  font-weight: bold;
}
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(class.source = "box-style")
options(scipen = 999)
require(cdx2cea)
require(readxl)
data("l_psa", package = "cdx2cea")
```

Target journal: Pharmacoeconomics

Type of paper: Practical application

# Introduction

Health economic (HE) models are routinely developed to inform health policy decisions such as including (new) healthcare interventions in insurance packages or restricting their use to specific subgroups[@drummond2015]. Validating health economic models is a crucial step within the process of developing health economic models to identify and correct potential errors[@radeva2020], to increase model credibility, and to reduce the risk of suboptimal policy decisions[@caro2014]. However, validation efforts on HE models are often not systematically performed nor reported[@abraham2025].

Validation of HE models is a multifaceted concept. The ISPOR-SMDM Modeling Good Research Practices Task Force defined different aspects of HE model validity: face validity, verification (also called internal validity), cross validity, and external validity of the model structure, inputs, and outputs [@eddy2012][@corroramos2024], but does not provide an easy-to-use tool to report validation efforts for each of these aspects. Following this initial definition of HE model validation, multiple generic tools have been developed to structure the (reporting of) the validation efforts performed on HE models. Well-known examples of such tools are AdviSHE tool, the TECH-VER checklist, and the CADTH’s Model Validation Tool to assist in the Conduct of Economic Evaluations[@vemer2016; @büyükkaramikli2019; @coyle2024]. Both AdviSHE and CADTH’s tool can be used to assess multiple aspects of validity, such as face validity and technical verification while TECH-VER solely focuses on the latter.

Besides the availability of these validation tools, the (increasing) use of script-based software, such as ‘R’ [@R_ref], to develop HE models[@jalal2017] facilitates the automated execution of systematic and generic validation tests and may be key to improve HE modelling validation practices . For instance, the `hesim` package[@hesim], *an R package for health economic simulation modeling and decision analysis*, contains a function to easily check the summary statistics of the parameter distributions used for the probabilistic analysis (PA). Similarly, the `darthpack` package, *an R package that showcases the Decision Analysis in R for Technologies in Health (DARTH) coding framework*, contains functions to assess the validity of the transition matrices and arrays used to populate health state transition models developed according to this framework[@alarid-escudero2019]. These examples show the feasibility of integrating simple validation tests during the development of script-based health economic models. Even though the nature of these validation tests is generic, their current implementation within coding frameworks for HE models may limit their use and usefulness beyond HE models developed outside these coding frameworks.

Hence, generic (R) software tools to easily and systematically validate HE models are required to improve HE model validation practices. Such tools may be particularly useful for developers who are new to script-based HE model development. Examples of such tools are the `assertHE` R package and the probabilistic check analysis dashboard (PACBOARD) [@smith2024; @pouwels2024]. `assertHE` depicts the network of functions defined and used within a HE model developed in R [@smith2024]. This package is useful for HE model developers and reviewers since one gets an overview of the backbone of the HE model. `assertHE` may further improve the communication of the workings of a developed HE models to a less technically-oriented audience. PACBOARD is a web-based dashboard which allows to systematically validate HE model input parameter and output values and allows to explore the relation between HE model input and outputs values. This dashboard is partly powered by the `pacheck` R package.

The `pacheck` R package offers a suite of functions aiming at validating HE models and exploring their workings using metamodelling techniques. Metamodels are statistical models, such as a linear regression model, fitted to the (probabilistic) inputs and outputs of a HE models [@degeling2018]. These metamodels allow to rapidly estimate the output of a HE models without requiring access to the source code, but also allow to assess the direction and magnitude of the relationships between inputs and outputs.

The validation tests included in `pacheck` are based on a published pragmatic literature review [@pouwels2024] . Most of the validation tests included in `pacheck` were identified in the TECH-VER checklist, emphasising their relevance for HE models. In contrast to with other R package dedicated to HE model development and validation, `pacheck` may be applied to inputs and outputs of HE models developed with other software `package` than R once the probabilistic inputs and outputs are loaded within the R session of interest. This broaden the scope and usefulness of `pacheck` to HE models developed in all possible software environments supporting the export of input and output data.

To support the uptake of the `pacheck` R package for model validation, clear examples of its validation are valuable to further increase the understanding of its functionality, input and output, and limitations. This paper therefore describes the functionalities of the `pacheck` R package using the probabilistic inputs and outputs of four published open-source HE models [@alarid-escudero2022; @incerti2019; @cusick2023; @sweeting2018], including example codes as tutorial for (novice) R-users.

The following section describes shortly the case studies. Next, the functionalities of `pacheck`, including annotated R code are showcased on one open-source model (the applications of `pacheck` functionalities on the other case studies are available in the Supplementary Material). We continue by reflecting on facilitators to apply `pacheck` in these different case studies, and conclude by discussing limitations and areas of further development of `pacheck`.

# Case studies

The functionalities of `pacheck` are illustrated using the open-source HE models included in the `cdx2cea` R package [ @alarid-escudero2022][@alarid-escudero2021], the `iviRA` R package[@iviRA; @incerti2019], the python-based chronic kidney disease screening cost-effectiveness analysis (CKDScreeningCEA) [@cusick2023], and the R-based discrete event simulation (DES) model for Abdominal Aortic Aneurysma (AAA) screening [@sweeting2018].

The `cdx2cea`package has been developed to perform a *"cost-effectiveness analysis (CEA) of testing average-risk Stage II colon cancer patients for the absence of CDX2 biomarker expression followed by adjuvant chemotherapy"* [@alarid-escudero2022]. The CEA supported by this package is performed using a cohort-based health state transition model. It compares two strategies: No CDX2 testing and CDX2 testing. The PA dataset, which is available in the `cdx2cea`package (`l_psa` object), contains the probabilistic HE model inputs and outputs of `r format(l_psa$n_sim, big.mark = ",")` iterations. The HE model contains `r length(l_psa$parnames)` input parameters included in the probabilistic analuysis and provides discounted quality-adjusted life years and discounted costs for both strategies. This case study is used to demonstrate that the functionalities of `pacheck` can be used on a HE model developed according to the DARTH coding framework.

The `iviRA` package is an individual-level health state transition model developed to assess the costs and health effects of different treatment sequences for rheumatoid arthritis. To demonstrate the use of `pacheck` we generated 1,000 probabilistic inputs and outputs simulating 100 individuals using the functions of the `iviRA` package. The (script used to generate these) inputs and outputs are available in the github repository of the `pacheck` package.

CKDScreeningCEA is a cohort-based health-state transition model, which estimates the health and economic impact of population-wide screening for chronic kidney disease (CKD) [@cusick2023]. The model was developed in Python and compares screening for CKD, starting at different ages and using different frequencies, with no screening. Probabilistic inputs of this health economic model are stored in an .xlsx file that can be directly imported in R using the `readxl` package [@readxl]. Alternatively, these inputs can be converted to .csv and / or .rda file before being imported in R. The probabilistic disaggregated outcomes of the model are also exported in an .xlsx file, which can be imported in R in a similar fashion as the inputs. For this illustration, we ran 1,000 probabilistic iterations of the model.

The AAA screening model is a DES model aiming at assessing the health and economic impact of screening for AAA in women [@sweeting2018]. All probabilistic inputs and outputs of this model are stored in a R list object. For this illustration, we ran 1,000 probabilistic iterations of the model containing 1,000 individuals.

All inputs and outputs datasets were generated using `r version$version.string` and are available in the `pacheck` package.

# Illustrations of functionalities

## Validating health economic model inputs and outputs

Functionalities of `pacheck` to assess the validity of the model input parameters concern the plausibility of the range in which model input parameters vary. For instance, `pacheck` contains simple validity assessments such as assessing whether utility values and probabilities remain between 0 and 1, and whether there are no occurrences of negative costs or other strictly positive parameters in each probabilistic iteration. `pacheck` also contains a function to assess and visualise whether two survival curves cross (`check_surv_mod` and `plot_surv_mod`). This assessment of plausibility is especially relevant when developing partitioned survival models, where a progression-free survival (PFS) curve should always result in lower probabilities than an overall survival curve (OS). Concerning the validation of model outputs, users are able to check whether total costs and effects are positive, whether discounted results are lower than undiscounted results, and whether the mean quality of life of both strategies is within the mean utility values used for the different health states of the model. All validation efforts are also performed for each iteration of the probabilistic analysis and the `pacheck` functions mentions in which iteration this occurred. `pacheck` can be used to visually assess the convergence of HE model outputs [@hatswell2018]. Box 1 further illustrates how to perform these validation efforts using different functions of `pacheck`.\

*Box 1: example R code of using `pacheck` to validate the health economic model inputs and outputs*

```{r valid_inputs_outputs, warning = FALSE}
# Install and load packages
# install.packages("devtools")
# devtools::install_github("feralaes/cdx2cea")
# devtools::install_github("DARTH-git/dampack")
# devtools::install_github("Xa4P/pacheck")
# install.packages("foreach")
require(cdx2cea)
require(dampack)
require(pacheck)

# Load & prepare dataframes for validation tests
data("l_psa", package = "cdx2cea")

# Inspect parameter values - limited to first 5 for the sake of brevity 
generate_sum_stats(l_psa$parameters[, 1:5])

# Test whether utility values are within the 0-1 
check_binary(c("u_Stg2", "u_Stg2Chemo", "u_Mets"),
             df = l_psa$parameters)

# Introduce utility values below 0 and above 1
df_inputs_error <- l_psa$parameters
df_inputs_error[c(4, 444, 754), "u_Stg2"] <- -1
df_inputs_error[c(3, 333, 681), "u_Stg2Chemo"] <- 99
df_inputs_error[c(5, 554, 153), "u_Mets"] <- -1
df_inputs_error[c(6, 146, 538), "u_Mets"] <- 99

# Check for erroneous utility values with pacheck function
check_binary(c("u_Stg2", "u_Stg2Chemo", "u_Mets"),
             df = df_inputs_error) 

# Perform multiple plausibility checks on a 'DARTH' psa object
data.frame(check_psa_darth(l_psa))

# Check convergence of a model output
## Plot moving average incremental effectiveness - per 100 iterations
l_psa$effectiveness$Incremental_QALYs <- l_psa$effectiveness$`CDX2 testing and FOLFOX if CDX2-negative` - l_psa$effectiveness$`No CDX2 testing and no FOLFOX`
plot_convergence(l_psa$effectiveness, 
                 param = "Incremental_QALYs", 
                 block_size = 100,
                 y_min = 0.03,
                 y_max = 0.04)
# minimum and maximum value set to prevent 'excessive' peaks
```

## Investigating the relationships between inputs and outputs

The `pacheck` package also contains various functions to investigate the relation between HE model inputs and outputs. For instance, the correlation matrix or plot between inputs and outputs can be calculated using the `generate_cor` function. Using the probabilistic inputs and outputs of the `cdx2cea`, one can see that the total costs of the intervention is negatively correlated with cancer mortality rate, `r_DieMets`,. This result seems logical since a higher probability of death would lead to shorter survival and thus lower costs. Linear regression metamodelling is also available through `pacheck` (using the `lm` function for the linear regression modelling). When estimating the incremental difference in QALYs between the intervention (CDX2 testing + adjuvant chemotherapy) and the comparator, we can see that an increase in utility values of stage II cancer with and without chemotherapy, `u_Stg2Chemo` and `u_Stg2`, leads to higher incremental QALYs while an increase in the utility value of the metastatic recurrent state, `u_Mets`, leads to lower incremental QALYs. These results are explainable when considering the model structure of this HE evaluation. In the intervention strategy, individuals spend more time in the stage II cancer with and without chemotherapy health state. Increasing the utility value of either `u_Stg2Chemo` (applied only during chemotherapy administration) or `u_Stg2` (applied after completion of chemotherapy treatment) therefore increases the number of QALYs accumulated in this health state in the intervention strategy versus the comparator strategy. The opposite applies when increasing the `u_Mets` utility value. Individuals spend less time in the metastatic health state in the intervention strategy compared to the the control strategy. Therefore, increasing this utility value decreases the incremental QALYs associated with the intervention versus the comparator strategy. A striking result of this metamodel, is its high (adjusted) R^2^ of 0.97. This can be explained by the likely linear relationships between the inputs and outputs of this simple HE model.

Metamodel validation can be performed in `pacheck` by using user-defined train-test split proportions or cross-validation using a user-defined number of folds. When using the train-test split method, the user needs to supply the partition of the data that will be used to train the model. In the example in Box 2, 75% of the data is used to train the model, and the remaining 25% are used to test the trained metamodel. During cross-validation, the complete dataset is divided in a number of folds, for instance 10. The metamodel is then trained on 9 of the 10 folds and validated in the remaining fold. This is repeated for each fold. The validation metrics of the metamodel are obtained by averaging the validation metrics of the metamodel in each fold (**See Appendix for an illustration of the implementation in `pacheck`**).

The high R^2^ value (0.97) and the position of the prediction versus observation dots in the calibration dots near the 45 degree line in the test set show that the metamodel may be deemed valid to estimate the incremental QALYs. When deemed valid, such metamodel may be used to perform sensitivity analyses, which is especially useful for computationally intensive HE models. In Box 2, we show that increasing the utility value for the metastatic state, `u_Mets`, to 0.5 (mean value in the probabilistic set was 0.25) is precited to increase the incremental QALYs by 0.001, from 0.035 in the base-case to 0.036.

As mentioned by Jalal et al., metamodel's parameters may be subject to scale effects. Jalal et al. therefore suggest to normalise HE model inputs before metamodelling, which facilitates their interpretation in relation to each other [@jalal2013]. Normalisation of inputs has been implemented in `pacheck` within the `lm_metamod` function. Box 2 illustrates how to use the functions of `pacheck` to investigate the relationships between HE model inputs and outputs.

*Box 2: Example R code of how to use the `pacheck` package to assess the relationships between inputs and outputs.*

```{r input_output, warning = FALSE}
# Generate correlation matrix between inputs and outputs
## Transform l_psa parameters and outcomes in one dataframe
df_psa_cdx2cea <- cbind(l_psa$parameters, 
                        l_psa$effectiveness, 
                        l_psa$cost)
## Calcualte correlations between first 5 inputs & total costs of the intervention
tbl_cor <- generate_cor(df_psa_cdx2cea[, c(1:5, ncol(df_psa_cdx2cea))]) 
## Show correlations between the 5 inputs & total costs of the intervention
tbl_cor[1:5, "CDX2 testing and FOLFOX if CDX2-negative"] 
## Display correlation matrix using 'tile' plot from ggplot2
generate_cor(df_psa_cdx2cea,
             figure = T) 

# Fit linear metamodel to predict the incremental QALYs between strategies
## Check relation between costs input parameters and incremental QALYs
v_c_vars <- l_psa$parnames[grep("^c_", l_psa$parnames)]
tbl_cor_iQALYs <- generate_cor(df_psa_cdx2cea[, c(v_c_vars, "Incremental_QALYs")],
             figure = F)
tbl_cor_iQALYs[v_c_vars, "Incremental_QALYs"] 
## No apparent correlation between costs input parameters and incremental QALYs

## Decision not to include costs input parameter in the linear metamodel
v_x_vars <- l_psa$parnames[grep("^c_", l_psa$parnames, invert = T)]
y_var <- names(l_psa$effectiveness)[3]
lm_metamod <- fit_lm_metamodel(y_var = y_var,
                               x_vars = v_x_vars, 
                               df = df_psa_cdx2cea)
# the names of the input parameters can be used instead of "v_x_vars" 

## Overview of estimated linear metamodel parameters
summary(lm_metamod$fit) 

# Normalise the inputs
lm_metamod_standardised <- fit_lm_metamodel(y_var = y_var,  
                                                     x_vars = v_x_vars,
                                                     df = df_psa_cdx2cea,
                                                     seed_num = 123,
                                                     standardise = TRUE)
summary(lm_metamod_standardised$fit) # provides an overview of estimated linear metamodel parameters

# Validation metamodel using the train-test approach
lm_metamod_valid <- fit_lm_metamodel(y_var = y_var,
                                     x_vars = v_x_vars,
                                     df = df_psa_cdx2cea,
                                     seed_num = 123,
                                     validation = "train_test_split",
                                     partition = 0.75) 
## in combination with the "train_test_split" approach, 
## the proportion of observation used to train the metamodel (`partition`) 
## should be provided. The remainder is used as validation check

lm_metamod_valid$stats_validation
lm_metamod_valid$calibration_plot

# Prediction metamodel
## Select input parameters
df_params <- data.frame(df_psa_cdx2cea[, v_x_vars])
## Increase the utility value of the metastatic state
df_params$u_Mets <- 0.5 
v_pred <- predict_metamodel(model = lm_metamod_valid,
                            inputs = df_params) 
## 'inputs' has to be a dataframe
c(average = round(mean(v_pred), 3),
  `percentile 2.5%` = unname(round(quantile(v_pred, probs = 0.025), 3)),
  `percentile 97.5%` = unname(round(quantile(v_pred, probs = 0.975), 3))) 
# averag  of 0.036 QALYs, base-case incremental QALYs was 0.035
```

Additional functionalities of `pacheck` are provided in the Supplemental Material, using the other three case studies **(REF to Supplemental Material)**.

# Facilitators to applying `pacheck` to different open-source models

Due to their open-source character, all selected HE models provided access to a full set of probabilistic inputs and outputs or we could easily generate such a set of probabilistic inputs and outputs. This allowed to quickly implement the functionalities of `pacheck` after running the model (in three of the four case studies) and limited manipulation of the format of the inputs and outputs. Only the `cdx2cea` provided a saved set of probabilistic inputs and outputs. Ensuring all inputs and outputs are saved, and ideally ordered as a table, facilitates the application of `pacheck`.

Each model had a clear README file, which facilitated navigation within the different file structure of the models. In addition, most of the inputs and outputs were described using explicit names. This was helpful to understand the potential natural bounds of these inputs and outputs. For instance, words and abbreviations such as 'rate', 'prob' (probabilities), 'HR' (hazard ratio), and 'cost' were used in multiple models. However, a clear definition of each input and output was most often missing. To ensure users have direct access to the meaning of the different variables represented by the input and output parameters of the model, we suggest that modellers add descriptive metadata to the probabilistic inputs and outputs files. Also, navigation within health economic models could be improved (especially for novice script-based software users) if all modellers would adhere to the same standard file structure and a coding style, such as the one proposed by the DARTH working group [@alarid-escudero2019].

# Discussion

The current paper presents functionalities of the `pacheck` R package and illustrates the practical use of these functionalities, with detailed R code, in four case studies.\
While we demonstrated that pacheck is easy to use and contains useful validation tests, the package clearly does not provide a complete list of validation tests for HE models. `Pacheck` is a package that requires regular updates and addition of validation tests over time to ensure it remains relevant for HE model developers and reviewers.\
From a technical point of view, `pacheck` is most likely not coded in the most efficient way. This is a design choice we made to ensure transparency of the implemented validation tests and to encourage external contributors to review the code base and to contribute to the further development of the package. The source code of `pacheck` is openly available on GitHub: [https://github.com/Xa4P/pacheck](#0){.uri}. External contributors can raise "Issues" concerning the package and propose new validation tests via "Issues" and "Pull requests". In future releases of `pacheck`, we aim to implement a larger number of validation tests and types of metamodel. We are also planning to test `pacheck` on the probabilistic inputs and outputs of "closed-source" HE models to demonstrate the usefulness of `pacheck` in such setting.\
Finally, `pacheck` focuses on the technical verification of HE model which is only a single aspect of validity. Hence, passing the validation tests included in `pacheck` should not create a feeling of false certainty concerning the validity of HE models. HE model developers and reviewers are therefore encouraged to use `pacheck` as a complement to other validation tools such as AdviSHE and CADTH's tool to report and review a broad range of validation aspects [@vemer2016, @coyle2024] .
