---
title: "Case study - iviRA model"
author: "Xavier G.L.V. Pouwels"
format: pdf
editor: visual
---

# Introduction

This vignette shows how to use the functionalities of `pacheck` on the `iviRA` health economic model.

# Loading packages and data

```{r}
#library(pacheck)
devtools::load_all()
data("l_iviRA_pa_params")
data("l_iviRA_out_summ_1")
data("l_iviRA_out_summ_2")
```

# Generating the probabilistic inputs and outputs

The source code to generate the model inputs and outputs of the `iviRA` cost-effectiveness model is available in the source code of the `pacheck` package on github (<https://github.com/Xa4P/pacheck/blob/master/analysis/03-iviRA_data_generation.R>).

# Preparing dataset for use in `pacheck`

The following code chunk prepares the inputs and outputs list of the `iviRA` package to use the `pacheck` tests. In this example, only a subset of all inputs is used for convenience.

```{r}
df_iviRA_pa <- data.frame(
  p_discount = l_iviRA_pa_params$tx.cost$discount,
  p_rebound = l_iviRA_pa_params$rebound,
  or_mort = exp(l_iviRA_pa_params$mort.logor),
  r_haq_prog = l_iviRA_pa_params$haq.lprog.tx,
  n_hosp_days = l_iviRA_pa_params$hosp.cost$hosp.days,
  c_hosp_pday = l_iviRA_pa_params$hosp.cost$cost.pday,
  c_si   = l_iviRA_pa_params$si.cost,
  c_prod_loss = l_iviRA_pa_params$prod.loss,
  u_model_wailoo = l_iviRA_pa_params$utility.wailoo,
  t_qaly_thx1 = l_iviRA_out_summ_1$means$qalys,
  t_qaly_d_thx1 = l_iviRA_out_summ_1$means$dqalys,
  t_cost_d_thx1 = l_iviRA_out_summ_1$means$dtot_cost,
  t_qaly_thx2 = l_iviRA_out_summ_2$means$qalys,
  t_qaly_d_thx2 = l_iviRA_out_summ_2$means$dqalys,
  t_cost_d_thx2 = l_iviRA_out_summ_2$means$dtot_cost
  )
df_iviRA_pa$inc_qaly <- df_iviRA_pa$t_qaly_d_thx1 - df_iviRA_pa$t_qaly_d_thx2
df_iviRA_pa$inc_cost <- df_iviRA_pa$t_cost_d_thx1 - df_iviRA_pa$t_cost_d_thx2
```

# Testing plausibility of model inputs and outputs

One way to test the plausibility of model inputs and outputs is simply to inspect the summary statistics of each variable. This can be done with the `pacheck` package using the `generate_sum_stats` function

```{r}
# Summary stats
generate_sum_stats(df = df_iviRA_pa)

# Specific checks
## Proportions & probabilities between 0 and 1
v_vars_discount <- grep("p_discount", names(df_iviRA_pa), value = TRUE) # returns variable names containing p_discount
check_binary(c(v_vars_discount, "p_rebound"), df = df_iviRA_pa)

## Costs, counts, OR positive
v_vars_n <- grep("n_", names(df_iviRA_pa), value = TRUE)
v_vars_c <- grep("c_", names(df_iviRA_pa), value = TRUE)
v_vars_c <- v_vars_c[-grep("inc_", v_vars_c)]
check_positive(c(v_vars_n, v_vars_c, "or_mort"), df = df_iviRA_pa)
do_check(df = df_iviRA_pa,
         v_vars = c(v_vars_n, v_vars_c, "or_mort"),
         check = ~ .x >= 0,
         label_check = "positive") # alternative

## Discounted results lower than undiscounted results
do_discount_check(df = df_iviRA_pa,
                  v_outcomes = c("t_qaly_thx1", "t_qaly_thx2"),
                  v_outcomes_d = c("t_qaly_d_thx1", "t_qaly_d_thx2"))
```

# Drawing health economic figures

`pacheck` also contains figures to draw standard figures for health economic evaluations, such as the cost-effectiveness plane (`plot_ce` function), the incremental cost-effectiveness plane (`plot_ice` function), the cost-effectiveness acceptability curves (`plot_ceac` function), and net benefit curves (`plot_nb` function). Both the `plot_ceac` and `plot_nb` functions require data frames containing specific variable names, which can respectively be obtained with the accompanying functions `calculate_ceac` and `calculate_nb`.

```{r}
plot_ce(df = df_iviRA_pa,
        e_int = "t_qaly_d_thx1",
        e_comp = "t_qaly_d_thx2",
        c_int = "t_cost_d_thx1",
        c_comp = "t_cost_d_thx2")

plot_ice(df = df_iviRA_pa,
        e_int = "t_qaly_d_thx1",
        e_comp = "t_qaly_d_thx2",
        c_int = "t_cost_d_thx1",
        c_comp = "t_cost_d_thx2")

df_ceac <- calculate_ceac(df = df_iviRA_pa,
        e_int = "t_qaly_d_thx1",
        e_comp = "t_qaly_d_thx2",
        c_int = "t_cost_d_thx1",
        c_comp = "t_cost_d_thx2")
plot_ceac(df = df_ceac,
          name_wtp = "WTP_threshold")

df_nb <- calculate_nb(df = df_iviRA_pa,
        e_int = "t_qaly_d_thx1",
        e_comp = "t_qaly_d_thx2",
        c_int = "t_cost_d_thx1",
        c_comp = "t_cost_d_thx2",
        wtp = 80000)
plot_nb(df = df_nb)
```

# Metamodelling

```{r}

```
