---
title: "Case study - Screening CKD CEA"
author: "Xavier G.L.V. Pouwels"
format: pdf
editor: visual
bibliography: references.bib
---

# Introduction

This vignette showcases how to use the functionalities of `pacheck` on a health economic model developed in Python focusing on the health economic impact on screening for Chronic Kidney Disease [@cusick2023].

# Loading packages and data

```{r}
#library(pacheck)
devtools::load_all()
set.seed(1234)
data("df_ckd_inputs")
data("df_ckd_results")
```

# Preparing dataset for use in `pacheck`

The following code chunk prepares the inputs and outputs list of the `iviRA` package to use the `pacheck` tests. In this example, only a subset of all inputs is used for convenience.

```{r, warning = FALSE}
df_ckd_results_wide <- df_ckd_results |>
  mutate(intervention = ifelse(intervention == "[0, 10, 20, 30, 40]", "Screening", intervention)) |>
  select(!c(full_index, Starting.age, dapa.costs, avg.on.dapa)) |>
  pivot_wider(names_from  = intervention,
              names_sep = "_",
              values_from = !intervention
              )
df_ckd_results_wide <- data.frame(apply(df_ckd_results_wide, 2, unlist))
df_ckd <- cbind(df_ckd_inputs,
                df_ckd_results_wide)
df_ckd$Unnamed..0 <- df_ckd$iteration <- NULL
```

# Testing plausibility of model inputs and outputs

One way to test the plausibility of model inputs and outputs is simply to inspect the summary statistics of each variable. This can be done with the `pacheck` package using the `generate_sum_stats` function. The correlation between (specific) inputs and outputs of the health economic model can be generated using the `generate_cor` function. This allows to investigate whether the relation between inputs and outputs

```{r}
# Summary stats
generate_sum_stats(df = df_ckd[, c(grep("Screening", names(df_ckd)),
                                   grep("Placebo", names(df_ckd)))]) 
## only for the outcomes for the sake of brievety

# Check correlations between inputs and outputs
##e.g. screening costs and higher ACE.costs should lead to higher screening costs and total costs
generate_cor(df = df_ckd,
             vars = c("screening_cost", "ACE.cost", "ace.costs_Screening", "ace.costs_Placebo", "costs_Screening", "costs_Placebo"))
## negative correlation between screening_cost and costs_Screening, 
## weak positive correlation between ACE.cost and total ACE costs 
## weak negative correlation with total costs of strategies 

# Specific checks
## utility values, specificity,  & sensitivity between 0 and 1
v_vars_u <- grep("_QALY", names(df_ckd), value = TRUE) # returns variable names
# containing _QALY
v_vars_u <- v_vars_u[-grep("disc", v_vars_u)]

check_binary(c(v_vars_u, "screening_sensitivity", "screening_specificity"), df = df_ckd)

## Costs, prevalence, HR positive
v_vars_prev <- grep("prev", names(df_ckd), value = TRUE)
v_vars_c <- grep("_cost", names(df_ckd), value = TRUE)
v_vars_hr <- grep(".HR", names(df_ckd), value = TRUE)
# v_vars_c <- v_vars_c[-grep("inc_", v_vars_c)]
check_positive(c(v_vars_prev, v_vars_c, v_vars_hr), df = df_ckd)

## Discounted results lower than undiscounted results
do_discount_check(
  df = df_ckd,
  v_outcomes = c(
    "QALY_Screening",
    "costs_Screening",
    "QALY_Placebo",
    "costs_Placebo"
  ),
  v_outcomes_d = c(
    "QALY.disc_Screening",
    "cost.disc_Screening",
    "QALY.disc_Placebo",
    "cost.disc_Placebo"
  )
)
```

# Drawing health economic figures

`pacheck` also contains figures to draw standard figures for health economic evaluations, such as the cost-effectiveness plane (`plot_ce` function), the incremental cost-effectiveness plane (`plot_ice` function), the cost-effectiveness acceptability curves (`plot_ceac` function), and net benefit curves (`plot_nb` function). Both the `plot_ceac` and `plot_nb` functions require data frames containing specific variable names, which can respectively be obtained with the accompanying functions `calculate_ceac` and `calculate_nb`.

```{r}
plot_ce(
  df = df_ckd,
  e_int = "QALY.disc_Screening",
  e_comp = "QALY.disc_Placebo",
  c_int = "cost.disc_Screening",
  c_comp = "cost.disc_Placebo"
)  
plot_ice(
  df = df_ckd,
  e_int = "QALY.disc_Screening",
  e_comp = "QALY.disc_Placebo",
  c_int = "cost.disc_Screening",
  c_comp = "cost.disc_Placebo"
)  
df_ceac <- calculate_ceac(
  df = df_ckd,
  e_int = "QALY.disc_Screening",
  e_comp = "QALY.disc_Placebo",
  c_int = "cost.disc_Screening",
  c_comp = "cost.disc_Placebo"
) 
plot_ceac(df = df_ceac, name_wtp = "WTP_threshold")  
df_nb <- calculate_nb(
  df = df_ckd,
  e_int = "QALY.disc_Screening",
  e_comp = "QALY.disc_Placebo",
  c_int = "cost.disc_Screening",
  c_comp = "cost.disc_Placebo",
  wtp = 80000
) 
plot_nb(df = df_nb)
```

# Reference
